---
title: "ED model output analyses"
output: html_notebook
          code_folding = true
        pdf_document
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse);
library(stringr);
```

Set outputs folder, specific id and get file
```{r}
outdir <- file.path("/home/luf74xx/Dokumente/model/EDoutputs")
outid <- file.path("beforeshutdown")
outraw <- as_tibble(read.table(file.path(outdir, outid, "orgsweekly.csv"), header = TRUE, sep = "\t"))
# very long file, since each organism gets a line
```

## Standard procedures
### Clean-up
```{r}
## parentheses from location ("()") and biomass ("()", "Dict", "=")+
loc <- gsub("[\\(|\\)]", "", outraw$location)
 ## split location in 3 columns
loc <- as.data.frame(matrix(unlist(str_split(loc, ",")),ncol =3,byrow = T))
names(loc) = c("xloc", "yloc", "floc")
## get vegetative and reproductive biomasses. When no reproductive biomass, adds 0 before reshaping 2 columns
masses = c()
for(plant in outraw$biomass){
  m <- unlist(str_extract_all(plant,"[0-9]{1,4}.[0-9]{1,3}")) # extracts both veg and repr
  if(length(m) < 2){
    m <- append(m, 0)
  }
  masses <- rbind(masses,m)
}
masses <- as.data.frame(masses, stringsAsFactors = FALSE); #there is probably a better way to convert into numeric
masses <- as.data.frame(sapply(masses, as.numeric));
names(masses) <-c("veg", "repr");
```

```{r}
# complete and clean table
outdata <- as_tibble(cbind(select(outraw,-one_of("location","biomass")),
                 loc,
                 masses))
rm(masses, loc, plant, m)
str(outdata)
head(outdata)
```

### Check output
The following should be used for every output analysis, to verify some general patterns

1. Check if reproduction is happening
```{r}
levels(outdata$reped)
```

2. Plot populations, weekly
  a. Extract biomass from raw and create separate data frame
  b. ggplot it seasonally
```{r}
weekabund <- outdata %>% group_by(week, sp) %>% summarize(abundance = n())

weekabund.plot <- ggplot(weekabund, aes(x = week, y = abundance, color = factor(sp))) 

weekabund.plot +
  geom_line() + #TODO differentiate species with colorgeom_line(color = "forestgreen") 
  labs(x = "Week", y = "Abundance (n individuals)",
       title = "Abundance variation",
       subtitle = "Species abundance per week")+
  theme(legend.position = "none")
```

3. Plot species-specific biomass, weekly
  a. Get species-specific biomass mean and variation weekly
  b. ggplot it seasonally
  c. `#TODO` so far,it is plotting the first year. For longer simulations, get mean from all years (since there is growth, get the mean of biomass uptake, instead of actual biomass) 
```{r}
weekmass <- outdata %>%
  group_by(week, sp) %>%
  summarise(veg.mean = mean(veg), veg.sd = sd(veg), repr.mean = mean(repr), repr.sd = sd(repr))

head(weekmass)

weekmass.plot <- ggplot(weekmass, aes(x=week, y= veg.mean, color = factor(sp)))

weekmass.plot +
  geom_line() +
  labs(x = "Week", y = "Mean vegetative biomass (g)",
       title = "Annual biomass variation",
       subtitle = "Mean vegetative biomass (+-sd) per species")+
  theme(legend.position = "none")
```
Same for reproductive biomass
```{r}
weekrepmass.plot <- ggplot(weekmass, aes(x=week, y= repr.mean, color = factor(sp)))

weekrepmass.plot + 
  geom_line() +
  labs(x = "Week", y = "Mean reproductive biomass (g)",
       title = "Annual reproductive biomass variation",
       subtitle = "Mean reproductive biomass (+-sd) per species")+
  theme(legend.position = "none")
```

4. Size (diameter) variation
```{r}
weeksize <- outdata %>%
  group_by(week, sp) %>%
  summarise(rad.mean = 3*mean(radius), rad.sd = 3*sd(radius)) #each cell is 3x3cm²

weeksize.plot <- ggplot(weeksize, aes(x=week, y=rad.mean, color = factor(sp)))

weeksize.plot +
  geom_line() +
  labs(x = "Week", y = "Mean radius (cm)",
       title = "Plant size variation (cm) ",
       subtitle = "Mean radius (+-sd) per species")+
  theme(legend.position = "none")
```

4. Biomass density over time
