---
title: "Single species simulation analysis"
output: html_notebook
          code_folding = true
        pdf_document
---
```{r, message=FALSE, warning=FALSE}
library(tidyverse);
library(stringr);
```

Set outputs folder, specific ID and get file
```{r}
outdir <- file.path("/home/ludmilla/Documents/uni_wuerzburg/phd_project/thesis/model/EDoutputs") #CCTBserver: /home/luf74xx/Dokumente/model/EDoutputs
outid <- file.path("singlesp1") #simID
outraw <- as_tibble(read.table(file.path(outdir, outid, "orgsweekly.csv"), header = TRUE, sep = "\t"))
# very long file, since each organism gets a line
head(outraw)
```

## Standard procedures
### Clean-up
```{r}
## parentheses from location ("()") and biomass ("()", "Dict", "=")+
loc <- gsub("[\\(|\\)]", "", outraw$location)
 ## split location in 3 columns
loc <- as.data.frame(matrix(unlist(str_split(loc, ",")),ncol =3,byrow = T))
names(loc) = c("xloc", "yloc", "floc")
## get vegetative and reproductive biomasses. When no reproductive biomass, adds 0 before reshaping 2 columns
masses = c()
for(plant in outraw$biomass){
  m <- unlist(str_extract_all(plant,"[0-9]{1,4}.[0-9]{1,3}")) # extracts both veg and repr
  if(length(m) < 2){
    m <- append(m, 0)
  }
  masses <- rbind(masses,m)
}
masses <- as.data.frame(masses, stringsAsFactors = FALSE); #there is probably a better way to convert into numeric
masses <- as.data.frame(sapply(masses, as.numeric));
names(masses) <-c("veg", "repr");

## Complete and clean table
outdata <- as_tibble(cbind(select(outraw,-one_of("location","biomass")),
                 loc,
                 masses))
rm(masses, loc, plant, m)
str(outdata)
head(outdata)
```

### Check output
The following should be used for every output analysis, to verify some general patterns

1. Verify input
*Temperature variation*
```{r}
temperature <- as_tibble(read.csv(file.path(indir,"envtimeseries_1999.csv"), header =TRUE))

annualtemp <- ggplot(temperature, aes(x=week, y =meantemp_ts))+
  geom_line()+
  labs(x = "Week", y = "Temperature (°C, monthly)",
       title = "Annual temperature variation")
annualtemp
```

*Species simulated*

```{r}
head(input)
```

1. Check if reproduction is happening
```{r}
levels(outdata$reped)
```

2. Plot populations, weekly
  a. Extract biomass from raw and create separate data frame
  b. ggplot it seasonally
```{r}
weekabund <- outdata %>% group_by(week, sp) %>% summarize(abundance = n())

weekabund.plot <- ggplot(weekabund, aes(x = week, y = abundance, color = factor(sp))) 

weekabund.plot +
  geom_line() + #TODO differentiate species with colorgeom_line(color = "forestgreen") 
  labs(x = "Week", y = "Abundance (n individuals)",
       title = "Abundance variation",
       subtitle = "Species abundance per week")+
  theme(legend.position = "none")
```

3. Plot individual vegetative biomass, weekly
  a. Get species-specific biomass mean and variation weekly
  b. ggplot it seasonally
  c. `#TODO` so far,it is plotting the first year. For longer simulations, get mean from all years (since there is growth, get the mean of biomass uptake, instead of actual biomass) 

Check individual variation
```{r}
biomass <- outdata %>%
  select(week,id,sp,veg,repr)

head(weekmass)

#Individual biomass variation
indweekmass.plot <- ggplot(biomass, aes(x=week, y= veg, color = factor(id)))
indweekmass.plot +
  geom_line() +
  labs(x = "Week", y = "Mean vegetative biomass (g)",
       title = "Annual biomass variation",
       subtitle = "Individual biomass")+
  theme(legend.position = "bottom")
```

Check variation at species-level
```{r}
# Species-specific biomass production
spproduction <- biomass%>%
  group_by(week, sp) %>%
  summarise(veg.mean = mean(veg), veg.sd = sd(veg), repr.mean = mean(repr), repr.sd = sd(repr))

spproduction.plot <- ggplot(spproduction, aes(x=week, y= veg.mean, color = factor(sp)))
spproduction.plot +
  geom_line() +
  labs(x = "Week", y = "Mean vegetative biomass (g)",
       title = "Annual biomass variation",
       subtitle = "Mean vegetative biomass (+-sd) per species")+
  theme(legend.position = "none")
```

Same for reproductive biomass
Individual variation
```{r}
indrepmass.plot <- ggplot(biomass, aes(x=week, y= repr, color = factor(id)))
indrepmass.plot + 
  geom_line() +
  labs(x = "Week", y = "Total reproductive biomass (g)",
       title = "Individual reproductive biomass variation")+
  theme(legend.position = "bottom")
```

Species-level variation
```{r}
sprepmass.plot <- ggplot(biomass, aes(x=week, y= repr, color = factor(sp)))
sprepmass.plot + 
  geom_line() +
  labs(x = "Week", y = "Mean reproductive biomass (g)",
       title = "Annual reproductive biomass variation",
       subtitle = "Mean reproductive biomass (+-sd) per species")+
  theme(legend.position = "none")
```

4. Size (diameter) variation
```{r}
indsize <- outdata %>%
  group_by(week, id) %>%
  summarise(rad.mean = 3*(radius+1)) #each cell is 3x3cm², plus central cell, to which the radius is added

indsize.plot <- ggplot(indsize, aes(x=week, y=rad.mean, color = factor(id)))

indsize.plot +
  geom_line() +
  labs(x = "Week", y = "Radius (cm)",
       title = "Plant size variation (cm) ",
       subtitle = "Individual radius")+
  theme(legend.position = "bottom")
```
