---
title: "Single species simulation analysis"
output: html_notebook
          code_folding = true
        pdf_document
---
```{r, message=FALSE, warning=FALSE}
library(tidyverse);
library(stringr);
```

# Single species, 1m² (Achillea millefolium)

## Standard procedures
Set outputs folder, specific ID and get file
```{r, include=FALSE}
#indir <- file.path("/home/ludmilla/Documents/uni_wuerzburg/phd_project/thesis/model/inputs") 
indir <- file.path("/home/luf74xx/Dokumente/model/inputs"); #CCTBserver
inid <- file.path("singlespinput1.csv") # check in simID file in output
input <- as_tibble(read.table(file.path(indir, inid), header = TRUE, sep = ","))
#outdir <- file.path("/home/ludmilla/Documents/uni_wuerzburg/phd_project/thesis/model/EDoutputs") #CCTBserver 
outdir <- file.path("/home/luf74xx/Dokumente/model/EDoutputs");
outid <- file.path("singlesp1"); #simID
outraw <- as_tibble(read.table(file.path(outdir, outid, "orgsweekly.csv"), header = TRUE, sep = "\t"));
# very long file, since each organism gets a line
```

Clean raw data
```{r, include=FALSE}
## parentheses from location ("()") and biomass ("()", "Dict", "=")+
loc <- gsub("[\\(|\\)]", "", outraw$location)
 ## split location in 3 columns
loc <- as.data.frame(matrix(unlist(str_split(loc, ",")),ncol =3,byrow = T))
names(loc) = c("xloc", "yloc", "floc")
## get vegetative and reproductive biomasses. When no reproductive biomass, adds 0 before reshaping 2 columns
masses = c()
for(plant in outraw$biomass){
  m <- unlist(str_extract_all(plant,"([0-9]{1,4}.[0-9]{1,7})")) # extracts both veg and repr
  if(length(m) < 2){
    m <- c(0,m)
  }
  masses <- rbind(masses,m)
}
masses <- as.data.frame(masses, stringsAsFactors = FALSE); #there is probably a better way to convert into numeric
masses <- as.data.frame(sapply(masses, as.numeric));
names(masses) <-c("repr", "veg");

## Complete and clean table
outdata <- as_tibble(cbind(select(outraw,-one_of("location","biomass")),
                 loc,
                 masses))
rm(masses, loc, plant, m)
```

## Check output

###Verify input behaviour
*Temperature variation* #TODO use landscape configuration output
*Ladscape configuration*
```{r}
temperature <- as_tibble(read.csv(file.path(indir,"envtimeseries_1999.csv"), header =TRUE))

ggplot(temperature, aes(x=week, y =meantemp_ts))+
  geom_line()+
  labs(x = "Week", y = "Temperature (°C, monthly)",
       title = "Annual temperature variation during 1999")
```

*Species simulated*
```{r}
View(input)
```

###Growth

1. Plot individual vegetative biomass, weekly
  a. Get species-specific biomass mean and variation weekly
  b. ggplot it seasonally
  c. `#TODO` so far,it is plotting the first year. For longer simulations, get mean from all years (since there is growth, get the mean of biomass uptake, instead of actual biomass) 

Check individual variation
```{r}
biomass <- outdata %>%
  select(week,id,sp,veg,repr)

#Individual biomass variation
indweekmass.plot <- ggplot(biomass, aes(x=week, y= veg, color = factor(id)))
indweekmass.plot +
  geom_line() +
  labs(x = "Week", y = "Mean vegetative biomass (g)",
       title = "Annual biomass variation",
       subtitle = "Individual biomass")+
  theme(legend.position = "bottom")
```
This linear growth is not due to the growth rate having always the same value, which would be an error. The growth rate is changing:

```{r}
production <- biomass %>%
  mutate(total = veg+repr) %>%
  group_by(week,id)

ind1 = diff(production$total[which(production$id == "p1-1")],lag = 1)
ind2 = diff(production$total[which(production$id == "p1-2")],lag = 1)
plot(ind1, type = "l", ylab = "Production (g)", xlab = "Week", main ="Individual biomass production")
lines(ind2, type = "l", col ="red")
legend("topleft", c("ind1","ind2"), fill =c("black","red"))
```

Check variation at species-level
```{r}
# Species-specific biomass production
spproduction <- biomass%>%
  group_by(week, sp) %>%
  summarise(veg.mean = mean(veg), veg.sd = sd(veg), repr.mean = mean(repr), repr.sd = sd(repr))

spproduction.plot <- ggplot(spproduction, aes(x=week, y= veg.mean, color = factor(sp)))
spproduction.plot +
  geom_line() +
  labs(x = "Week", y = "Mean vegetative biomass (g)",
       title = "Annual biomass variation",
       subtitle = "Mean vegetative biomass (+-sd) per species")+
  theme(legend.position = "none")
```

Same for reproductive biomass
Individual variation
```{r}
indrepmass.plot <- ggplot(biomass, aes(x=week, y= repr, color = factor(id)))
indrepmass.plot + 
  geom_line() +
  labs(x = "Week", y = "Total reproductive biomass (g)",
       title = "Individual reproductive biomass variation")+
  theme(legend.position = "bottom")
```

Species-level variation
```{r}
sprepmass.plot <- ggplot(biomass, aes(x=week, y= repr, color = factor(sp)))
sprepmass.plot + 
  geom_line() +
  labs(x = "Week", y = "Mean reproductive biomass (g)",
       title = "Annual reproductive biomass variation",
       subtitle = "Mean reproductive biomass (+-sd) per species")+
  theme(legend.position = "none")
```

4. Size (diameter) variation
```{r}
indsize <- outdata %>%
  group_by(week, id) %>%
  summarise(rad.mean = 3*(radius+1)) #each cell is 3x3cm², plus central cell, to which the radius is added

indsize.plot <- ggplot(indsize, aes(x=week, y=rad.mean, color = factor(id)))

indsize.plot +
  geom_line() +
  labs(x = "Week", y = "Radius (cm)",
       title = "Plant size variation (cm) ",
       subtitle = "Individual radius")+
  theme(legend.position = "bottom")
```

###Reproduction

1. Check if reproduction is happening
```{r}
levels(outdata$reped)
```

2. Plot populations, weekly
```{r}
weekabund <- outdata %>% group_by(week, sp) %>% summarize(abundance = n())

weekabund.plot <- ggplot(weekabund, aes(x = week, y = abundance, color = factor(sp))) 

weekabund.plot +
  geom_line() + #TODO differentiate species with colorgeom_line(color = "forestgreen") 
  labs(x = "Week", y = "Abundance (n individuals)",
       title = "Abundance variation",
       subtitle = "Species abundance per week")+
  theme(legend.position = "none")
```

## Unrestricted growth and seed number 
```{r}
fgoutid <- file.path("singlespfullgrowth"); #simID
fgoutraw <- as_tibble(read.table(file.path(outdir, fgoutid, "orgsweekly.csv"), header = TRUE, sep = "\t"));
# very long file, since each organism gets a line
```

Clean raw data
```{r, include=FALSE}
## parentheses from location ("()") and biomass ("()", "Dict", "=")+
loc <- gsub("[\\(|\\)]", "", fgoutraw$location)
 ## split location in 3 columns
loc <- as.data.frame(matrix(unlist(str_split(loc, ",")),ncol =3,byrow = T))
names(loc) = c("xloc", "yloc", "floc")
## get vegetative and reproductive biomasses. When no reproductive biomass, adds 0 before reshaping 2 columns
masses = c()
for(plant in fgoutraw$biomass){
  m <- unlist(str_extract_all(plant,"([0-9]{1,4}.[0-9]{1,7})")) # extracts both veg and repr
  if(length(m) < 2){
    m <- c(0,m)
  }
  masses <- rbind(masses,m)
}
masses <- as.data.frame(masses, stringsAsFactors = FALSE); #there is probably a better way to convert into numeric
masses <- as.data.frame(sapply(masses, as.numeric));
names(masses) <-c("repr", "veg");

## Complete and clean table
fgoutdata <- as_tibble(cbind(select(fgoutraw,-one_of("location","biomass")),
                 loc,
                 masses))
rm(masses, loc, plant, m)
```

###Growth

1. Plot individual vegetative biomass, weekly
  a. Get species-specific biomass mean and variation weekly
  b. ggplot it seasonally
  c. `#TODO` so far,it is plotting the first year. For longer simulations, get mean from all years (since there is growth, get the mean of biomass uptake, instead of actual biomass) 

Check individual variation
```{r}
fgbiomass <- fgoutdata %>%
  select(week,id,stage,sp,veg,repr) #there was reproduction, and seed mass is very different

#Individual adult biomass variation
ggplot(filter(fgbiomass, stage == "a"), aes(x=week, y= veg, color = factor(id))) +
  geom_line() +
  labs(x = "Week", y = "Mean vegetative biomass (g)",
       title = "Annual biomass variation",
       subtitle = "Individual biomass")+
  theme(legend.position = "bottom")
```
This linear growth is not due to the growth rate having always the same value, which would be an error. The growth rate is changing:

```{r}
production <- fgbiomass %>%
  mutate(total = veg+repr) %>%
  group_by(week,id)

ind1 = diff(production$total[which(production$id == "p1-1")],lag = 1)
ind2 = diff(production$total[which(production$id == "p1-2")],lag = 1)
plot(ind1, type = "l", ylab = "Production (g)", xlab = "Week", main ="Individual biomass production")
lines(ind2, type = "l", col ="red")
legend("topleft", c("ind1","ind2"), fill =c("black","red"))
```

Check variation at species-level
```{r}
# Species-specific biomass production
spproduction <- fgbiomass%>%
  group_by(week, sp) %>%
  summarise(veg.mean = mean(veg), veg.sd = sd(veg), repr.mean = mean(repr), repr.sd = sd(repr))

spproduction.plot <- ggplot(spproduction, aes(x=week, y= veg.mean, color = factor(sp)))
spproduction.plot +
  geom_line() +
  labs(x = "Week", y = "Mean vegetative biomass (g)",
       title = "Annual biomass variation",
       subtitle = "Mean vegetative biomass (+-sd) per species")+
  theme(legend.position = "none")
```

Same for reproductive biomass
Individual variation
```{r}
indrepmass.plot <- ggplot(fgbiomass, aes(x=week, y= repr, color = factor(id)))
indrepmass.plot + 
  geom_line() +
  labs(x = "Week", y = "Total reproductive biomass (g)",
       title = "Individual reproductive biomass variation")+
  theme(legend.position = "bottom")
```

Species-level variation
```{r}
sprepmass.plot <- ggplot(fgbiomass, aes(x=week, y= repr, color = factor(sp)))
sprepmass.plot + 
  geom_line() +
  labs(x = "Week", y = "Mean reproductive biomass (g)",
       title = "Annual reproductive biomass variation",
       subtitle = "Mean reproductive biomass (+-sd) per species")+
  theme(legend.position = "none")
```

4. Size (diameter) variation
```{r}
indsize <- outdata %>%
  group_by(week, id) %>%
  summarise(rad.mean = 3*(radius+1)) #each cell is 3x3cm², plus central cell, to which the radius is added

indsize.plot <- ggplot(indsize, aes(x=week, y=rad.mean, color = factor(id)))

indsize.plot +
  geom_line() +
  labs(x = "Week", y = "Radius (cm)",
       title = "Plant size variation (cm) ",
       subtitle = "Individual radius")+
  theme(legend.position = "bottom")
```

###Reproduction

1. Check if reproduction is happening
```{r}
levels(fgoutdata$reped)
```

2. Plot populations, weekly
```{r}
weekabund <- fgoutdata %>% group_by(week, sp) %>% summarize(abundance = n())

weekabund.plot <- ggplot(weekabund, aes(x = week, y = abundance, color = factor(sp))) 

weekabund.plot +
  geom_line() + #TODO differentiate species with colorgeom_line(color = "forestgreen") 
  labs(x = "Week", y = "Abundance (n individuals)",
       title = "Abundance variation",
       subtitle = "Species abundance per week")+
  theme(legend.position = "none")
```